Exploring Design Space for Differentially Private Bayesian Inference

### 2019-01-13 ICML2019/main.tex
pass on sec 2

\begin{abstract}
% Several approaches have been explored in order to make Bayesian inference
% differentially private.
% For instance, \cite{dimitrakakis2014robust} and
% \cite{wang2015privacy} proved that, under specific conditions of the likelihood function and the prior distribution,
% sampling from the posterior distribution, is already differentially private. \cite{zhang2016differential} and \cite{foulds2016theory}
% designed differentially private (d.p.) mechanisms by adding laplacian noise
% to the numeric parameters of the posterior distributions. 
% Accuracy, is usally (\cite{??}) measured by providing upper bounds, holding w.h.p, on the
% $\ell_1$ or $\ell_\infty$-distance between the numeric parameters
% of the released posterior distribution and the actual real ones of posterior distribution.
We explore the design space for differentially private Bayesian
inference.  We focused on two discrete models for parametric Bayesian
inference i.e., the Beta-Binomial and the Dirichlet-multinomial
models. We propose a variation of the exponential mechanism that uses
a scoring function based on a metric over probailistic measures, i.e.,
the Hellinger distance. We compare it with other approaches based on
the perturbation of the parameters of the posterior distribution.  We
measure the accuracy using the Hellinger distance over the posterior
distribution released and the real one as opposed to measuring the
$\ell_1$ or $\ell_{\infty}$ distance on the parameters.  We finally
present an experimental analysis on the accuracy of these mechanisms.
\end{abstract}
 % TODO: replace with your keywords

@@

\noindent \textbf{Main contributions.}
\begin{itemize}
	\item We designed a differentially private Bayesian inference mechanism based on the standard exponential mechanism.

@@

\section{Mechanism Proposition}
\label{sec_mechs}
\subsection{Laplace Mechanism Family}

\subsubsection{Using $\ell_1$ Norm Metric}
  We firstly present the fully explored mechanism in Laplace family -- Laplace mechanism. This mechanism ($\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$) simply adds Laplace noise with post-processing to the posterior distribution parameters, producing an output: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$ with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into $[0,n]$ to make sure the noised posterior is valid, as shown in Mechanism \ref{mech:lapmech}.\\
  Since we are adding noise directly to the parameter, the sensitivity should be calibrated to $\ell_1$ of two Beta distributions:
  $
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
  $,
  i.e., proportional to the dimensionality.

  \begin{algorithm}
  \caption{$\lapmech$ in Beta-binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad {\bf let} $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad {\bf let} $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \ENSURE $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

  In Dirichlet-multinomial model, $\lapmech$ produces output: $ \dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0},$
  where $\eta_i \thicksim \lap{0}{\frac{k - 1}{\epsilon}}$.
  We will use $ \lapmech(\dataobs) = \betad(\tilde{\alpha}, \tilde{\beta})$ or $\dirichlet{\tilde{\valpha}}$ for short.

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The $\lapmech$ presented above can be improved in our Bayesian inference model, by adding noise scaled to smaller sensitivity in this improved Laplace mechanism $\ilapmech$.\\
 In terms of two adjacent data sets $\dataobs, \dataobs'$, their posterior distributions by Bayesian inference -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

>>>>

\begin{abstract}
% Several approaches have been explored in order to make Bayesian inference
% differentially private.
% For instance, \cite{dimitrakakis2014robust} and
% \cite{wang2015privacy} proved that, under specific conditions of the likelihood function and the prior distribution,
% sampling from the posterior distribution, is already differentially private. \cite{zhang2016differential} and \cite{foulds2016theory}
% designed differentially private (d.p.) mechanisms by adding laplacian noise
% to the numeric parameters of the posterior distributions. 
% Accuracy, is usally (\cite{??}) measured by providing upper bounds, holding w.h.p, on the
% $\ell_1$ or $\ell_\infty$-distance between the numeric parameters
% of the released posterior distribution and the actual real ones of posterior distribution.
In this work we explore the design space for differentially private Bayesian
inference.  We focus on two discrete models for parametric Bayesian
inference i.e., the Beta-Binomial and the Dirichlet-Multinomial
models. We propose a variation of the exponential mechanism that uses
a scoring function based on a metric over probabilistic measures, i.e.,
the Hellinger distance. We compare it with other approaches based on
the perturbation of the parameters of the posterior distribution.  We
measure the accuracy using the Hellinger distance over the posterior
distribution released and the real one as opposed to measuring the
$\ell_1$ or $\ell_{\infty}$ distance on the parameters.  We finally
present an experimental analysis on the accuracy of these mechanisms.
\end{abstract}
 % TODO: replace with your keywords

@@

\noindent \textbf{Main contributions.}
\begin{itemize}
	\item We design a new differentially private Bayesian inference mechanism based on the standard exponential mechanism.

@@

\section{Private Mechanisms}
\label{sec_mechs}
\subsection{Calibrating noise w.r.t. $\ell_1$ norm}
We start off by presenting a mechanism based on the Laplace primitive.
The mechanism, denoted by {\color{red}{i think this macro has too many arguments}}($\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$),
simply adds Laplace noise to the parameters of the posterior distributions. The mechanisms also performs a
post-processing phase by {\color{red}{flooring}} and  truncating the noisy parameters so that they belong to the right interval.
In the case of the Beta-Binomial model the output produced is: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$
with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into
$[0,n]$. This mechanism is described by the pesudocode in Figure \ref{mech:lapmech}.\\
Since we are adding noise directly to the parameter, the sensitivity should be calibrated to $\ell_1$ of two Beta distributions:
$
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
$,
i.e., proportional to the dimensionality.
  \begin{algorithm}
  \caption{$\lapmech$ in Beta-binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad {\bf let} $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad {\bf let} $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \ENSURE $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

  In Dirichlet-Multinomial model, $\lapmech$ produces output:
  $\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0},$
  where $\eta_i \thicksim \lap{0}{\frac{k - 1}{\epsilon}}$.
  We will use $ \lapmech(\dataobs) = \betad(\tilde{\alpha}, \tilde{\beta})$ or $\dirichlet{\tilde{\valpha}}$ for short.

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The accuracy of the previous $\lapmech$ can be improved, by adding noise scaled to smaller sensitivity.
 In terms of two adjacent data sets $\dataobs, \dataobs'$, the posterior distributions that they induce -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}
